---
title: "Homework 2"
author: "Alexandra Thursland"
date: "8/30/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

### Task 3

Write a function that takes as its inputs that data you simulated (or any data of the same type) and a sequence of θ values of length 1000 and produces Likelihood values based on the Binomial Likelihood. Plot your sequence and its corresponding Likelihood function.
The likelihood function is given below. Since this is a probability and is only valid over the interval from [0, 1] we generate a sequence over that interval of length 1000.
You have a rough sketch of what you should do for this part of the assignment. Try this out in lab on your own.

```{r}
### Bernoulli LH Function ### # Input: obs.data, theta
# Output: bernoulli likelihood
### Plot LH for a grid of theta values ### # Create the grid #
# Store the LH values
# Create the Plot
```

### Task 4

Write a function that takes as its inputs prior parameters a and b for the Beta-Bernoulli model and the observed data, and produces the posterior parameters you need for the model. *Generate and print* the posterior parameters for a non-informative prior i.e. (a,b) = (1,1) and for an informative case (a,b) = (3,1)}.

### Task 5

Create two plots, one for the informative and one for the non-informative case to show the posterior distribution and superimpose the prior distributions on each along with the likelihood. What do you see? Remember to turn the y-axis ticks off since superimposing may make the scale non-sense.

## Question 2

(20 points total) The Exponential-Gamma Model We write X ∼ Exp(θ) to indicate that X has the Exponential distribution, that is, its p.d.f. is
p(x|θ) = Exp(x|θ) = θ exp(−θx)1(x > 0).
The Exponential distribution has some special properties that make it a good model for certain applications. It has been used to model the time between events (such as neuron spikes, website hits, neutrinos captured in a detector), extreme values such as maximum daily rainfall over a period of one year, or the amount of time until a product fails (lightbulbs are a standard example).
Suppose you have data x1, . . . , xn which you are modeling as i.i.d. obser- vations from an Exponential distribution, and suppose that your prior is θ ∼ Gamma(a, b), that is,
ba
p(θ) = Gamma(θ|a, b) = Γ(a) θa−1 exp(−bθ)1(θ > 0).

### a

(5) Derive the formula for the posterior density, p(θ|x1:n). Give the form of the posterior in terms of one of the most common distribu- tions (Bernoulli, Beta, Exponential, or Gamma).

### b

(5) Why is the posterior distribution a proper density or probability distribution function?

### c

(5) Now, suppose you are measuring the number of seconds between lightning strikes during a storm, your prior is Gamma(0.1, 1.0), and your data is
(x1 , . . . , x8 ) = (20.9, 69.7, 3.6, 21.8, 21.4, 0.4, 6.7, 10.0).
Plot the prior and posterior p.d.f.s. (Be sure to make your plots on
a scale that allows you to clearly see the important features.)

### d 

(5) Give a specific example of an application where an Exponential model would be reasonable. Give an example where an Exponential model would NOT be appropriate, and explain why.

## Question 3

(40 points total) Priors, Posteriors, Predictive Distributions (Hoff, 3.9) An unknown quantity Y has a Galenshore(a, θ) distribution if its density is given by
p(y) = 2 θ2ay2a−1e−θ2y2 Γ(a)
for y > 0,θ > 0,a > 0. Assume for now that a is known. For this density, E[Y]= Γ(a+1/2)
   and
θΓ(a) E[Y2]= a.
θ2

### a

(10) Identify a class of conjugate prior densities for θ. Plot a few members of this class of densities.

### b

(5) Let Y1, . . . , Yn ∼ Galenshore(a, θ). Find the posterior distribu-
tion of θ | y1:n using a prior from your conjugate class.

### c

(10) Write down
p(θa | y1:n)
p(θb | y1:n) and simplify. Identify a sufficient statistic.

### d

(5) Determine E[θ | y1:n].

### e

(10) Show that the form of the posterior predictive density
2y2a−1Γ(an + a + c) n+1
(d2 + 􏰀 y2)an+c i
p(yn+1 | y1:n) = Γ(a)Γ(an + c)
(d2 + 􏰀 yi2 + yn2+1)(an+a+c) .

